On The Benefits of Magnitude Estimation Relevance Assessments for Information Retrieval Evaluation

### 2016-09-15 response.tex

In the related work section, please make precise how the TOIS
submission differs from the earlier SIGIR 2015 paper.

\comment{A paragraph has been added at the end of Section 2.1 to do this.}

RQ2 is an important additional ingredient of this extended paper.
Section 6 is devoted to answering this question. Please explicitly
announce that that is what you are going to do. And please provide
answers to the question at the end of section 6. We are now left
to discover for ourselves what (if any) answer you were able to
produce for RQ2.

\comment{We have added Section 6.4 which summarises the section.}

@@

\noindent ...

\section{Reviewer 2}

I like the paper and the problem that it addresses: magnitude
estimation for information retrieval and the use of crowd sourcing
for this purpose. The authors claim that their findings are direct
implications for IR evaluation. This is probably true, but I think
the authors can do a better job of articulating them. I recommend
that each result section 5-8 explicitly (1) starts with one of the
four research questions, (2) answers it, (3) explains what the
ramifications of the answers are for IR evaluation.

\comment{
This has been done for Section 6 (RQ2) by adding Section 6.4, 
Section~7 (RQ3) by adding Section~7.4, and
Section~8 (RQ4) by adding Section~8.1.
We feel RQ1 is reasonably covered in Section~5.
}

@@

Some more specific comments:

Section 2.3: good addition. You could consider adding a brief
discussion on the value of crowd worker disagreement. There is work
by Lora Aroyo and Chris Welty that you should probably relate to
in section 2.3

\comment {Thanks for pointing this out, we have added a paragraph in Section 2.3.}

@@

Figure 7 has no legends (neither x nor y).

\comment {Curious: there are axis labels, and the symbols are explained
in the caption. Perhaps a printing error on the reviewer's side?}

What are the practical implications of the statement at the very
end of section 6, about the number of required workers? And what
does it say about setups where a smaller number of workers has been
used?

\comment {We have added Section 6.4 to draw this section together.}

@@

Section 7 coincides with the old section 5. 
Figs 10, 11, 12 need legends.

\comment {Again, this seems fine on our copies.}

It would be good to explain the differences in absolute
numbers between Table IV in the TOIS submission and Table 2 in the
SIGIR paper. And why these differences do not matter.

\comment {We have added two sentences at the end of the first paragraph in Section~7.2 
explaining the difference. On closer inspection, we also noticed that the new methodology did make a
difference, and have altered the text in the final two paragraphs of Section~7.3 accordingly.}

@@

Section 8: I am not sure I am happy with the uncertain ending of
this section. What does the disagreement with Kanoulas and Aslam's
earlier work mean for our "ideal" evaluation setup?

\comment{We have added Section~8.2 Summary to clarify this.}

@@

Section 9: the addition of a "limitations" section is an excellent
idea.

Appendix A: I like this addition. A screen dump of the instruction
interface would have been helpful.

\comment{The exact instructions that appeared on the screen are already given in the appendix, and a screen
dump would just format them in an HTML/www style, rather than text. As such, we 
don't think a screen dump adds anything here. However, even if the reviewer is insistent, it turns 
out that CrowdFlower has undergone an `upgrade' since we
ran the experiments, and, sadly, the exact screen shot is no longer available anyway.
}
\end{document}

### 2016-09-15 10_conc.tex
OK - all comments addressed. Tidied up SM's. "Fixed" section 7. Still some sm comments in (abstract?). Needs careful read (esp section 7).

This study is the first to investigate document level relevance
judgments using magnitude estimation, at scale.
A much-cited benefit of magnitude estimation is that it is a ratio
scaling technique.
Indeed, workers were instructed to assign relevance scores as ratios
to those assigned to previous documents.
However, some documents in the collected data lead to ratios that have
a large number of significant figures.
For example, in one unit, the first document was given a score of 54,
and the second 76, implying that the worker thought that the second
document was 1.407 times more relevant than the first.
This seems like an unusually high amount of precision for such a task,
and it may be that the worker was choosing numbers without closely
following the ratio requirement.
Intuitively, one might think that it would be possible to distinguish
the relevance of two documents at a ratio of one significant figure
(for example, ``2 times as relevant'', or ``100 times as relevant'');
or maybe even two significant figures (``21 times as relevant'', or
''140 times as relevant''); but it seems very unlikely that this could
be done to three significant figures or more without extensive training
and effort.
We can safely assume that our crowdsourced workers were not trained in
relevance assessment, nor spent more than a minute or two on each
document (mean time to assess 78 seconds, standard deviation 86
seconds).

>>>>

This study is the first to investigate document level relevance
judgments using magnitude estimation, at scale.
A much-cited benefit of magnitude estimation is that it is a ratio
scaling technique.
Indeed, workers were instructed to assign relevance scores as ratios
to those assigned to previous documents.
However, some documents in the collected data lead to ratios that have
a large number of significant figures.
For example, in one unit, the first document was given a score of 54,
and the second 76, implying that the worker thought that the second
document was 1.407 times more relevant than the first.
This seems like an unusually high amount of precision for such a task,
and it may be that the worker was choosing numbers without closely
following the ratio requirement.
Intuitively, one might think that it would be possible to distinguish
the relevance of two documents at a ratio of one significant figure
(for example, ``2 times as relevant'', or ``100 times as relevant'');
or maybe even two significant figures (``21 times as relevant'', or
``140 times as relevant''); but it seems very unlikely that this could
be done to three significant figures or more without extensive training
and effort.
We can safely assume that our crowdsourced workers were not trained in
relevance assessment, nor spent more than a minute or two on each
document (mean time to assess 78 seconds, standard deviation 86
seconds).

### 2016-09-15 6_results_RQ2.tex
OK - all comments addressed. Tidied up SM's. "Fixed" section 7. Still some sm comments in (abstract?). Needs careful read (esp section 7).

\sm{Let us wrap up an overall answer to  RQ\ref{item:rq4}. As we anticipated, using ME with crowdsourcing could potentially raise several concerns, including: subjectiveness of the scale, normalization issues, lack of proper training, potential internal and/or external disagreement, and overall cost in terms of the number of redundant workers needed.}
\sm{On the basis of the results presented so far we can conclude that, at least with our setting and our quality checks in place (see items (a)--(d) in Section~\ref{sec:quality-checks}), we can be reassured that magnitude estimation relevance scores gathered by means of crowdsourcing are reliable. Also, the cost is competitive, as in our experiment it is in the same range of the cost of collecting ordinal scale judgments, and there are clues that it can be lowered by collecting fewer judgments.}

>>>>

\textcolor{red}{ 
Using ME with crowdsourcing could potentially raise
several concerns, including: 
subjectiveness of the relevance scale; 
difficulty in normalizing disparate scales from different workers; 
lack of proper training of workers;
potential internal and/or external disagreement between workers;
and high overall cost as redundant workers needed to ensure quality.
On the basis of the results presented so far we can conclude that, at
least with our setting and our quality checks in place (see items
(a)--(d) in Section~\ref{sec:quality-checks}), we can be reassured that
magnitude estimation relevance scores gathered by means of
crowdsourcing are reliable, thus answering RQ\ref{item:rq4}.
Also, the cost to collect the ME judgements in our experiments 
is comparable to that of collecting ordinal scale judgements, 
and there are hints that this cost can be lowered by collecting fewer judgments.
}

### 2016-09-15 7_results_RQ3.tex
OK - all comments addressed. Tidied up SM's. "Fixed" section 7. Still some sm comments in (abstract?). Needs careful read (esp section 7).

Given that the agreement between our magnitude scores, TREC judgments
and Sormunen's judgments is not perfect, it is reasonable to expect
that relative system effectiveness orderings computed with each of
these as a basis may differ.
We first examine the correlation between system orderings using
magnitude estimation scores and TREC relevance as gain values, as for
both these judgment sets we have nearly complete coverage of the top 10
documents of all runs submitted to TREC-8, for our 18 judged topics.
% In the conference version of this paper~\cite{ME-SIGIR15} we showed the
% system scores (mean over all 18 topics) using median magnitude
% estimation scores per document (x-axis), and binary TREC categories
% (y-axis).
% This gave a result similar to Figure~\ref{fig:runs-rerank-median3best}.
% Unjudged documents were assumed to be irrelevant, as is common practice
% when using TREC data.
% \aht{What do you think of this almost self-referential style?}
% 
% Here we make a small alteration to the aggregation of the ME scores to
% get a single score for a document.
Rather than taking the median of all scores assigned to a document, we
take the median of scores taken from the three units that have document
orderings that have the highest agreement with the TREC orderings, or
Sormunen orderings, respectively.
In effect, the scores derived from units with high internal agreement as
computed in Section~\ref{sec:internal-agreement}, with ties broken in favor
of units with the least unjudged documents.
By aggregating scores over units that have a high agreement with the
original TREC orderings, any differences in system rankings we see should be more
attributable to the scale of the gain values, rather than perturbations in
document orderings given by different gain values.
\textcolor{red}{Note that this is a slightly less confounded methodology 
than was used in our previous paper~\cite{ME-SIGIR15}, which simply took the median 
of all units, and accounts for the 
differences between the figures and tables in this section and 
that previous paper.}

@@

%%%\sm{I've been thinking and rethinking over this. Some remarks:
%%%  \begin{itemize}
%%%  \item I think we felt short with this analysis. 
%%%    By selecting individual judges we sort of lose the benefit of the
%%%    crowd: it is rather accepted, if not well known, that
%%%    \emph{redundant} judgments need to be collected, in any
%%%    situation.  So in place of using individual scores I'd prefer to
%%%    use the median score of a proper subset of all 10 workers. 
%%%    (I'm not sure it will make a difference since we're taking several
%%%    samples, but still I guess it will be different)
%%%  \item Another reason for the low tau correlations below could be
%%%    that individual workers did not simply disagree on the magnitudes
%%%    and the scales, but also on the ranking. 
%%%    But we're neglecting this.
%%%    A better, or at least alternative, way of doing these experiments
%%%    would be to select only the workers that have a pairwise agreement
%%%    high enough with the ``official'' judges, so that any disagreement
%%%    can truly be ascribed to different scales. 
%%%    This is consistent with what Falk wrote in his last note above: we
%%%    select the workers with high pairwise agreement and we investigate
%%%    their variability.
%%%  \item All this needs to be re-thought after the results of
%%%    Section~\ref{sec:how-many-workers}. 
%%%  \item So in summary: take Section~\ref{sec:how-many-workers};
%%%    consider worker subsets rather than individual workers; select
%%%    only the workers with pairwise agreement high enough. Hm?
%%%  \end{itemize}
%%%  \aht{OK, i agree there is a confound that and this can be imrpoved.
%%%I think we leave 6.3 out of this paper, and that makes this section 
%%%easier too. I will do (14 Dec).}
%%%}

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=.6\linewidth]{figs/sysRank_scatter_TREC_med3bestME.pdf}
  \caption{System scores using TREC categories (y-axis) 
    and magnitudes (x-axis) as gains in 
    the nDCG@10 and ERR@10 metrics. There is one dot for each of the 129 systems 
    participating in the TREC-8 ad hoc track, with red dots showing the top-set (see text).
    Magnitudes used are the median of the three units per document that most agree with TREC
    document rankings.
    Kendall's $\tau$ is shown. 
  \label{fig:runs-rerank-median3best}
  }
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=.6\linewidth]{figs/sysRank_scatter_FAKE_medME3bestME.pdf}
  \caption{As for Figure~\ref{fig:runs-rerank-median3best} but using 
           Sormunen categories (y-axis) and simulated runs as described in the text.
  \label{fig:runs-rerank-median3best-fake}
  }
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=.7\linewidth]{figs/metrics_resample.pdf}
  \caption{Kendall's $\tau$ between system rankings obtained using gains from 
    the judgments and metrics indicated on the x-axis, where magnitudes are randomly
    sampled for each document 1000 times. Compare with $\tau$ in
Figures~\ref{fig:runs-rerank-median3best} and~\ref{fig:runs-rerank-median3best-fake}.
  \label{fig:runs-rerank-resample}
  }
\end{figure}

An alternate explanation for the low $\tau$ values is that gains set as
magnitudes are generally much higher than the gains set by Sormunen's
categories.
From Figure~\ref{fig:ME-distribution} it is apparent that magnitudes
are generally in the range of 2 to 15 (although some are as large as
100 or 1000), whereas using categories as gains the values are 0, 1, 2
or 3. However, because both nDCG and ERR are normalized, this scale
effect is nullified to a degree.
For example, multiplying gains by a constant has no effect on nDCG, and
even altering the category scores using a small exponential or additive
constant has little effect.
Table~\ref{tab:scaling-cats} shows this Kendall's $\tau$ for our 12
topics on the simulated systems using various mappings of Sormunen
categories to gain values, and a mapping of those categories.
The final two rows hint that using gain values from a broad scale,
while preserving document orderings, may alter system rankings,
particularly with nDCG@10, compared with simply using the category
values as gains.
The final row perhaps best represents our magnitude scores, and so goes
partway to explaining the low $\tau$ values in the previous section.

\begin{table}[tp]
\begin{center}
\tbl{Correlation (Kendall's $\tau$) and overlap of top-set
between system orderings when using Sormunen
categories as gain ($G_i=C_i$), and other mappings
between categories and gains
on the simulated systems,
when using the nDCG@10 and ERR@10 metrics.
\label{tab:scaling-cats}
}{%
\begin{tabular}{lcc r cc}
\toprule
  & \multicolumn{2}{c}{Kendall's $\tau$} 
  && 
    \multicolumn{2}{c}{Top-set overlap}  \\
 Mapping                  & nDCG@10 & ERR@10     && nDCG@10 & ERR@10 \\
\midrule
% $G_i=C_i$               &  1.000 & 1.000       &&  \ 32\%  & \ 68\% \\
% $G_i=$ med-best-3       &  0.147 & 0.060       &&  100\%  & 100\% \\
 $G_i= 100\times C_i$     &  1.000 & 0.698       &&  100\%  & 100\% \\
 $G_i= 1+C_i$             &  0.989 & 0.969       &&  \ 98\%  & \ 95\% \\
 $G_i= 10+C_i$            &  0.961 & 0.939       &&  \ 95\%  & \ 91\% \\
 $G_i= 100+C_i$           &  0.957 & 0.939       &&  \ 95\%  & 100\% \\
 $G_i= 10^{1+C_i}$        &  0.720 & 0.698       &&  \ 66\%  & 100\% \\
 $G_i= 2^{1+C_i}$         &  0.870 & 0.702       &&  \ 42\%  & \ 73\% \\
\bottomrule
\end{tabular}%
}
\end{center}
\end{table}

% \sm{It is a perhaps bit strange that our lower median ME scores is
% not zero but 2 (Figure~\ref{fig:ME-distribution} and the text above).
% I know zero was not an option but I don't like that because not
% relevant should be zero or something like that.
% Maybe using as aggregation function the geometric mean in place of the
% median would help?
% Maybe we can simply subtract 2?!
% } 
%\aht{agreed - that is a little strange. I checked that all is kosher (14 Dec) and yes - it is
%the agregation function == median that is creating the floor of 2. Let's leave it as median in
%this paper though so that it is similar to the SIGIR paper (same reviewers, etc).}
% \fns{Also it perhaps doesn't matter too much, these ME scores are
% relative and perception-based.}
% \aht{You ``gain'' something from a N. Rel.}
% \aht{Note that in the SIGIR paper I had used Pearson, not Kendall - whoops.
% Also - the top set is strange for ERR} 
% 
% 
% \aht{deleted whole section on narrow and wide. It seems to me that the results were
% indicative at best, and just a distraction.}
% 
% \sm{but we should be adding things (+20\%), not removing them :-)}
% 
% \sm{I'm re-adding it, but I'm not convinced. Anyway, everything needs
%   to be double checked, the figure references are broken and tau
%   values in the table are wrong}

One way to examine this further is to split our data into
\emph{narrow} units, those whose $\mbox{\hkh}/\mbox{\nkn}$ ratios are
less than 5 (the median value), and \emph{wide} units, where the
ratios $\mbox{\hkh}/\mbox{\nkn}$ are 5 or more.

Table~\ref{tab:thin-med-thick} shows the correlation (Kendall's
$\tau$) between system orderings, and overlap of runs in the top set,
when using TREC and Sormunen judgments, and magnitude estimation
scores when the median is obtained from units that used a narrow scale,
a wide scale, or from all units (note that the $\tau$ values in the All
column match those in Figures~\ref{fig:runs-rerank-median3best}
and~\ref{fig:runs-rerank-median3best-fake}).
% \sm{false, taus are different now}
% \ref{fig:runs-rerank-median}
% and~\ref{fig:runs-rerank-median-fake}).
In each case the score for a topic-document pair is taken as 
the median of the 3 units that most agree with TREC or Sormunen 
and that are either narrow, wide or any as appropriate.
When there was no narrow or wide judgment for a topic-document pair
(all units containing that topic-document pair had
$\mbox{\hkh}/\mbox{\nkn}$ above or below the median), the median of the three best units was used.
This occurred in 12\% of TREC topic-document pairs, and 7\% of Sormunen topic-document pairs.
\textcolor{green}{PANIC!!!!
System orderings based on narrow or wide units show greater differences
(lower $\tau$ values) than when considering the full data set.
This is consistent with the results in Table~\ref{tab:scaling-cats}.
The overlap of the top set is generally less than 100\% for the
narrow/wide data sets, indicating that some perturbation of the set of
``equivalent'' top systems is occurring, but the pattern is not as clear as
for $\tau$ values.
This may be due to the relatively small number of topics that is being
considered, especially for the simulated systems, and more topics are
required before a definitive conclusion about the effect of narrow and
wide units on the top set can be made.
}

\begin{table}[t]
\begin{center}
\tbl{Correlation (Kendall's $\tau$) and overlap of the top set 
between system 
orderings when median document magnitude scores are based on 
narrow, wide or all units. 
\label{tab:thin-med-thick}
}{%
\begin{tabular}{rccc}
\toprule
\multicolumn{1}{l}{Source of gains} & Narrow & All & Wide \\
\midrule
\multicolumn{1}{l}{TREC, $\tau$} \\ \cmidrule(l){1-1} %cline{1-1}
nDCG@10 & 0.679 & 0.677 & 0.744 \\ 
ERR@10  & 0.361 & 0.222 & 0.247 \\ 
\addlinespace
\multicolumn{1}{l}{TREC, top set} \\ \cmidrule(l){1-1} 
nDCG@10 & 63.64\% & 81.82\% & 72.73\% \\ 
ERR@10  & 20.00\% & 20.00\% & 60.00\% \\

\addlinespace
\multicolumn{1}{l}{Sormunen-Simulated, $\tau$} \\  \cmidrule(l){1-1} 
nDCG@10 & 0.036 & 0.147 & 0.382 \\ 
ERR@10  & 0.017 & 0.060 & 0.323 \\ 
\addlinespace
\multicolumn{1}{l}{Sormunen-Simulated, top set} \\ \cmidrule(l){1-1} 
nDCG@10 & \ 9.76\% & 31.71\% & \ 68.29\% \\ 
ERR@10  & 68.18\% & 68.18\% & 100.00\% \\

@@

Overall, there are clear differences in system effectiveness orderings
as measured using Kendall's $\tau$,
when using narrow and wide units.
In particular, it appears that using the median of narrow units, or of
all units, leads to higher correlations with existing measurement
techniques, compared to when using wide units.
We can therefore infer that current evaluations using nDCG@10 and ERR@10
do not assume wide units as their underlying user model.
This is important because it shows that nearly half of our
participants are not behaving according to the user model assumed by
current evaluation techniques.
Figure~\ref{fig:hn-ratio} shows the full distribution of the ``width''
($\mbox{\hkh}/\mbox{\nkn}$ ratio) of units rounded to the nearest 5 
(with bars containing less than 4 units omitted).

>>>>

Given that the agreement between our magnitude scores, TREC judgments
and Sormunen's judgments is not perfect, it is reasonable to expect
that relative system effectiveness orderings computed with each of
these as a basis may differ.
We first examine the correlation between system orderings using
magnitude estimation scores and TREC relevance as gain values, as for
both these judgment sets we have nearly complete coverage of the top 10
documents of all runs submitted to TREC-8, for our 18 judged topics.
% In the conference version of this paper~\cite{ME-SIGIR15} we showed the
% system scores (mean over all 18 topics) using median magnitude
% estimation scores per document (x-axis), and binary TREC categories
% (y-axis).
% This gave a result similar to Figure~\ref{fig:runs-rerank-median3best}.
% Unjudged documents were assumed to be irrelevant, as is common practice
% when using TREC data.
% \aht{What do you think of this almost self-referential style?}
% 
% Here we make a small alteration to the aggregation of the ME scores to
% get a single score for a document.
Rather than taking the median of all scores assigned to a document, we
take the median of scores taken from the three units that have document
orderings that have the highest agreement with the TREC orderings, or
Sormunen orderings, respectively.
In effect, the scores derived from units with high internal agreement as
computed in Section~\ref{sec:internal-agreement}, with ties broken in favor
of units with the least unjudged documents.
By aggregating scores over units that have a high agreement with the
original TREC orderings, any differences in system rankings we see should be more
attributable to the scale of the gain values, rather than perturbations in
document orderings given by different gain values.
\textcolor{red}{Note that this is a slightly less confounded methodology 
than was used in our previous paper~\cite{ME-SIGIR15}, which simply took the median 
of all units, even if the unit wildly disagreed with the ordering of documents
given by the categorical judgements. This revised methodology accounts for the 
differences between the figures and tables in this section and 
in the previous paper.}

@@

%%%\sm{I've been thinking and rethinking over this. Some remarks:
%%%  \begin{itemize}
%%%  \item I think we felt short with this analysis. 
%%%    By selecting individual judges we sort of lose the benefit of the
%%%    crowd: it is rather accepted, if not well known, that
%%%    \emph{redundant} judgments need to be collected, in any
%%%    situation.  So in place of using individual scores I'd prefer to
%%%    use the median score of a proper subset of all 10 workers. 
%%%    (I'm not sure it will make a difference since we're taking several
%%%    samples, but still I guess it will be different)
%%%  \item Another reason for the low tau correlations below could be
%%%    that individual workers did not simply disagree on the magnitudes
%%%    and the scales, but also on the ranking. 
%%%    But we're neglecting this.
%%%    A better, or at least alternative, way of doing these experiments
%%%    would be to select only the workers that have a pairwise agreement
%%%    high enough with the ``official'' judges, so that any disagreement
%%%    can truly be ascribed to different scales. 
%%%    This is consistent with what Falk wrote in his last note above: we
%%%    select the workers with high pairwise agreement and we investigate
%%%    their variability.
%%%  \item All this needs to be re-thought after the results of
%%%    Section~\ref{sec:how-many-workers}. 
%%%  \item So in summary: take Section~\ref{sec:how-many-workers};
%%%    consider worker subsets rather than individual workers; select
%%%    only the workers with pairwise agreement high enough. Hm?
%%%  \end{itemize}
%%%  \aht{OK, i agree there is a confound that and this can be improved.
%%%I think we leave 6.3 out of this paper, and that makes this section 
%%%easier too. I will do (14 Dec).}
%%%}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.6\linewidth]{figs/sysRank_scatter_TREC_med3bestME.pdf}
  \caption{System scores using TREC categories (y-axis) 
    and magnitudes (x-axis) as gains in 
    the nDCG@10 and ERR@10 metrics. There is one dot for each of the 129 systems 
    participating in the TREC-8 ad hoc track, with red dots showing the top-set (see text).
    Magnitudes used are the median of the three units per document that most agree with TREC
    document rankings.
    Kendall's $\tau$ is shown. 
  \label{fig:runs-rerank-median3best}
  }
\end{figure}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.6\linewidth]{figs/sysRank_scatter_FAKE_medME3bestME.pdf}
  \caption{As for Figure~\ref{fig:runs-rerank-median3best} but using 
           Sormunen categories (y-axis) and simulated runs as described in the text.
  \label{fig:runs-rerank-median3best-fake}
  }
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.7\linewidth]{figs/metrics_resample.pdf}
  \caption{Kendall's $\tau$ between system rankings obtained using gains from 
    the judgments and metrics indicated on the x-axis, where magnitudes are randomly
    sampled for each document 1000 times. Compare with $\tau$ in
Figures~\ref{fig:runs-rerank-median3best} and~\ref{fig:runs-rerank-median3best-fake}.
  \label{fig:runs-rerank-resample}
  }
\end{figure}

An alternate explanation for the low $\tau$ values is that gains set as
magnitudes are generally much higher than the gains set by Sormunen's
categories.
From Figure~\ref{fig:ME-distribution} it is apparent that magnitudes
are generally in the range of 2 to 15 (although some are as large as
100 or 1000), whereas using categories as gains the values are 0, 1, 2
or 3. However, because both nDCG and ERR are normalized, this scale
effect is nullified to a degree.
For example, multiplying gains by a constant has no effect on nDCG, and
even altering the category scores using a small exponential or additive
constant has little effect.
Table~\ref{tab:scaling-cats} shows the effect of using different methods 
for converting the categorical values ($C_i$) from Sormunen into gain values
($G_i$) for computing nDCG and ERR
using our 12 topics on the simulated systems.
Kendall's $\tau$ is computed between the system rankings given using 
the categorical values as gains ($C_i=G_i$), and the other methods, 
hence the top row has perfect correlation.
The final two rows hint that using gain values from a broad scale,
while preserving document orderings, may alter system rankings,
particularly with nDCG@10, compared with simply using the category
values as gains.
The final row perhaps best represents our magnitude scores, and so goes
part way to explaining the low $\tau$ values in the previous section.
It does not, however, explain all of the changes in system rankings: 
it still seems that our ME judgments are different in more than scale.

\begin{table}[tp]
\begin{center}
\tbl{Correlation (Kendall's $\tau$) and overlap of top-set
between system orderings when using Sormunen
categories as gain ($G_i=C_i$), and other mappings
between categories and gains
on the simulated systems,
when using the nDCG@10 and ERR@10 metrics.
\label{tab:scaling-cats}
}{%
\begin{tabular}{lcc r cc}
\toprule
  & \multicolumn{2}{c}{Kendall's $\tau$} 
  && 
    \multicolumn{2}{c}{Top-set overlap}  \\
 Mapping                  & nDCG@10 & ERR@10     && nDCG@10 & ERR@10 \\
\midrule
$G_i=C_i$               &  1.000 & 1.000       &&  100\%  & 100\% \\
% $G_i=$ med-best-3       &  0.147 & 0.060       &&  100\%  & 100\% \\
 $G_i= 100\times C_i$     &  1.000 & 0.698       &&  100\%  & 100\% \\
 $G_i= 1+C_i$             &  0.989 & 0.969       &&  \ 98\%  & \ 95\% \\
 $G_i= 10+C_i$            &  0.961 & 0.939       &&  \ 95\%  & \ 91\% \\
 $G_i= 100+C_i$           &  0.957 & 0.939       &&  \ 95\%  & 100\% \\
 $G_i= 10^{1+C_i}$        &  0.720 & 0.698       &&  \ 66\%  & 100\% \\
 $G_i= 2^{1+C_i}$         &  0.870 & 0.702       &&  \ 42\%  & \ 73\% \\
\bottomrule
\end{tabular}%
}
\end{center}
\end{table}

% \sm{It is a perhaps bit strange that our lower median ME scores is
% not zero but 2 (Figure~\ref{fig:ME-distribution} and the text above).
% I know zero was not an option but I don't like that because not
% relevant should be zero or something like that.
% Maybe using as aggregation function the geometric mean in place of the
% median would help?
% Maybe we can simply subtract 2?!
% } 
%\aht{agreed - that is a little strange. I checked that all is kosher (14 Dec) and yes - it is
%the aggregation function == median that is creating the floor of 2. Let's leave it as median in
%this paper though so that it is similar to the SIGIR paper (same reviewers, etc).}
% \fns{Also it perhaps doesn't matter too much, these ME scores are
% relative and perception-based.}
% \aht{You ``gain'' something from a N. Rel.}
% \aht{Note that in the SIGIR paper I had used Pearson, not Kendall - whoops.
% Also - the top set is strange for ERR} 
% 
% 
% \aht{deleted whole section on narrow and wide. It seems to me that the results were
% indicative at best, and just a distraction.}
% 
% \sm{but we should be adding things (+20\%), not removing them :-)}
% 
% \sm{I'm re-adding it, but I'm not convinced. Anyway, everything needs
%   to be double checked, the figure references are broken and tau
%   values in the table are wrong}

One way to examine this further is to split our data into
\emph{narrow} units, those whose $\mbox{\hkh}/\mbox{\nkn}$ ratios are
less than 5 (the median value), and \emph{wide} units, where the
ratios $\mbox{\hkh}/\mbox{\nkn}$ are 5 or more.
Figure~\ref{fig:hn-ratio} shows a histogram of the number of units with each ratio.
This creates a set of judgments with a scale similar to Sormunen (the narrow units), and 
a set of judgments that are over a larger range (wide units).

Table~\ref{tab:thin-med-thick} shows the correlation (Kendall's
$\tau$) between system orderings, and overlap of runs in the top set,
when using TREC and Sormunen judgments, and magnitude estimation
scores when the median is obtained from units that used a narrow scale,
a wide scale, or from all units (note that the $\tau$ values in the All
column match those in Figures~\ref{fig:runs-rerank-median3best}
and~\ref{fig:runs-rerank-median3best-fake}).
% \sm{false, taus are different now}
% \ref{fig:runs-rerank-median}
% and~\ref{fig:runs-rerank-median-fake}).
In each case the score for a topic-document pair is taken as 
the median of the 3 units that most agree with TREC or Sormunen 
and that are either narrow, wide or any as appropriate.
When there was no narrow or wide judgment for a topic-document pair
(all units containing that topic-document pair had
$\mbox{\hkh}/\mbox{\nkn}$ above or below the median), the median of the three best units was used.
This occurred in 12\% of TREC topic-document pairs, and 7\% of Sormunen topic-document pairs.

\begin{table}[t]
\begin{center}
\tbl{Correlation (Kendall's $\tau$) and overlap of the top set 
between system 
orderings when median document magnitude scores are based on 
narrow, wide or all units. 
\label{tab:thin-med-thick}
}{%
\begin{tabular}{rccc}
\toprule
\multicolumn{1}{l}{Source of gains} & Narrow & Wide & All \\
\midrule
\multicolumn{1}{l}{TREC, $\tau$} \\ \cmidrule(l){1-1} %cline{1-1}
nDCG@10 & 0.679  & 0.744 & 0.677\\ 
ERR@10  & 0.361  & 0.247 & 0.222\\ 
\addlinespace
\multicolumn{1}{l}{TREC, top set} \\ \cmidrule(l){1-1} 
nDCG@10 & 63.64\% & 72.73\% & 81.82\% \\ 
ERR@10  & 20.00\% & 60.00\% & 20.00\% \\

\addlinespace
\multicolumn{1}{l}{Sormunen-Simulated, $\tau$} \\  \cmidrule(l){1-1} 
nDCG@10 & 0.036 & 0.382 & 0.147  \\ 
ERR@10  & 0.017 & 0.323 & 0.060  \\ 
\addlinespace
\multicolumn{1}{l}{Sormunen-Simulated, top set} \\ \cmidrule(l){1-1} 
nDCG@10 & \ 9.76\% & \ 68.29\% & 31.71\%  \\ 
ERR@10  & 68.18\%  & 100.00\%  & 68.18\% \\

@@

\textcolor{green}{
Generally the correlations between the system rankings based on 
categorical and ME gain values are higher when the wide judgements
are used as opposed to the narrow judgements.
This contradicts our expectations from 
Table~\ref{tab:scaling-cats}, which indicated that
a larger scale of gains led to a decrease in correlations.
This confirms that a simple argument based on the scale 
of judgements is not sufficient in explaining different system orders, 
although the number of topics used here is small (12).
We examine this further in the next section.
}

%Overall, there are clear differences in system effectiveness orderings
%as measured using Kendall's $\tau$,
%when using narrow and wide units.
%In particular, it appears that using the median of narrow units, or of
%all units, leads to higher correlations with existing measurement
%techniques, compared to when using wide units.
%We can therefore infer that current evaluations using nDCG@10 and ERR@10
%do not assume wide units as their underlying user model.
%This is important because it shows that nearly half of our
%participants are not behaving according to the user model assumed by
%current evaluation techniques.

\subsection{Summary}
\label{sec:summary-7}

\textcolor{red}{
In answering RQ\ref{item:rq2}, we have shown that there are clear
differences in system rankings using nDCG and ERR metrics on this data
when using ME scores as gain values compared with using ordinal
categories as gain values.
Depending on the judgements and the metric used, correlations between
system orderings using ME and the categorical judgements, as measured
with Kendall's $\tau$, varied between about 0.8 and -0.2. 
At first this might seem like a simple scaling problem, where 
ME judgements are over a wider range than categorical judgments.
While this is true, a much more likely explanation is that there is
substantial individual variance in not only how people perceive the
relevance of an individual document, but also in how they perceive the
ratio of relevance of one document to another.
}

### 2016-09-15 8_results_RQ4.tex
OK - all comments addressed. Tidied up SM's. "Fixed" section 7. Still some sm comments in (abstract?). Needs careful read (esp section 7).

Kanoulas and Aslam~\cite{KanAsl03} recommend a ratio of 1.1 for the
gain values of relevant to highly relevant documents to maximize the
effectiveness of a system comparison experiment.
Using magnitudes directly as gain values gives a median of 1.00, 1.00,
and 1.14 in the three boxes in Figure~\ref{fig:ka}, which happens to be
close to their recommendation.
However, again, we see a wide variation from this median, and mean
values of the ratios are 48, 22 and 24 respectively, further suggesting
that there may not be a single gain setting that is suitably
representative across different user perceptions and tasks.
% This hints that using median aggregation might be best for maximising the statistical efficiency of system
% comparisons using magnitudes as gain values.
% \aht{Really - does this last sentence even make logical senes? median of ratios matches, so use median
% aggregation???}

@@

>>>>

Kanoulas and Aslam~\cite{KanAsl03} recommend a ratio of 1.1 for the
gain values of relevant to highly relevant documents to maximize the
effectiveness of a system comparison experiment.
Using magnitudes directly as gain values gives a median of 1.00, 1.00,
and 1.14 in the three boxes in Figure~\ref{fig:ka}, which happens to be
close to their recommendation.
However, again, we see a wide variation from this median, and mean
values of the ratios are 48, 22 and 24 respectively, further suggesting
that there may not be a single gain setting that is suitably
representative across different user perceptions and tasks.
% This hints that using median aggregation might be best for maximizing the statistical efficiency of system
% comparisons using magnitudes as gain values.
% \aht{Really - does this last sentence even make logical senes? median of ratios matches, so use median
% aggregation???}

@@

\subsection{Summary}

\textcolor{red}{
In the previous section, and in this section, we have seen not only 
that relevance values for a single document vary amongst users, but
the ratio of relevance between two documents also varies.
Thus, in answer to RQ\ref{item:rq3}, the selection of a single 
gain formulation as either
a linear or exponential mapping of categorical judgements
is guaranteed to misrepresent at least half of our users 
in this data set.
Similarly, choosing a single gain formulation to maximize statistical 
efficiency of system comparisons will also misrepresent a large 
part of the user population.
}
